---
title: "02 Working Paper"
author: "Nate Breznau & Hung H.V. Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

## R Markdown

```{r setup}

source("R/build_causal_matrix.R")
source("R/build_formula_matrix.R")
source("R/keep_minimal.R")
```

## Generate Model Multiverse

### Input variables

Generate causal matrix

```{r models}
inputs <- list(nodes=c("a","b","c","d"), timing=c(0,-1,-2,-2),
              types=c("otc","test","ctr","ctr"))

causal_matrix <- build_causal_matrix(inputs)
```

### Derive formulas

Generate formula matrix

```{r formula_matrix}

formula_matrix <- build_formula_matrix(causal_matrix)
```

### Add extra plots

```{r xmodels}
formula_matrix <- as.data.frame(formula_matrix)

# Model 3
formula_matrix[(nrow(formula_matrix)+1),] <- c("X4 ~ Y + Xtest,Y ~ X2 + Xtest,Xtest ~ X2 + X1", as.numeric(nrow(formula_matrix)+1))

# Model 5
formula_matrix[(nrow(formula_matrix)+1),] <- c("Y ~ X2 + Xtest,X3 ~ X2,Xtest ~ X3 + X2 + X1", as.numeric(nrow(formula_matrix)+1))

```


### Derive plots

#### Generate DAG formulas

```{r viz}
coords <- list(
    x = c(Y = 1, Xtest = 0, X1 = -1, X2 = -1, X3 = -0.5, X4 = 1.5),
    y = c(Y = 0, Xtest = 0, X1 = -1, X2 = 1, X3 = 1, X4 = -1)
)

additional_args <- list(
    exposure="Xtest",
    outcome="Y",
    coords = coords
)

full_matrix <- formula_matrix
full_matrix[8,1] <- "Y ~ X2 + Xtest, Xtest ~ X2, X1 ~ X1" # this still doesn't work, the plotting drops the X1
full_matrix_dags <- list()

# note that dagitty uses its own type of object, this means they need to be stored separately in a list
for (f in 1:nrow(formula_matrix)) {
    fvector <- strsplit(unlist(formula_matrix[f,1]), ",")[[1]]
    # create dag object syntax
    dag <- do.call(dagify, c(lapply(fvector, as.formula), additional_args))
    model <- full_matrix$model[f]
    full_matrix_dags[[model]] <- dag
    # extract adjustment sets
    mas <- adjustmentSets(dag, exposure = "Xtest", outcome = "Y", effect = "direct")
    mas <- ifelse(mas == "list()", "none", capture.output(mas))
    full_matrix$mas[f] <- paste(unlist(mas), collapse = ",")
}

## X1 is at the corner, clearly not removed. Perhaps because you've set the coordinates and some of the coordinates are outside of the graph?
test_dag <- dagify(Y ~ X2 + Xtest, Xtest ~ X2, X1 ~ X1)
plot(graphLayout(test_dag))
```

#### DAG Plots

Model 1 = 22 (Version B = 21)
Model 2 = 1
Model 3 = 33
Model 4 = 8 
Model 5 = 34
Model 6 = 18

```{r plots}

# until build formula is updated X1 does not appear in the plot if it is not connected to anything

for (l in 1:length(full_matrix_dags)){
    p <- paste0("p",l)
    assign(p, 
           plot(full_matrix_dags[[l]]))        
    text(0.5,-1, paste0("MAS =",full_matrix$mas[l]))
    text(0,1, paste0("Model ",l))
}


```
## Finding Multi MAS DAGs

Using our multiverse generator we search for DAGs that will not easily follow our classification rules

### Generate Models

```{r searching_dags}
inputs <- list(nodes=c("a","b","c","d", "e"), timing=c(0,-1,-2,-3,-3),
              types=c("otc","test","ctr","ctr", "ctr"))

causal_matrix2 <- build_causal_matrix(inputs)

formula_matrix2 <- build_formula_matrix(causal_matrix2)



## Add MAS
```


### Find MAS
```{r dagmassearch}
coords <- list(
    x = c(Y = 1, Xtest = 0, X1 = -1, X2 = -2, X3 = -2),
    y = c(Y = 0, Xtest = 0, X1 = 0.5, X2 = 1, X3 = -1)
)

additional_args <- list(
    exposure="Xtest",
    outcome="Y",
    coords = coords
)

full_matrix2 <- formula_matrix2
full_matrix_dags2 <- list()

# note that dagitty uses its own type of object, this means they need to be stored separately in a list
for (f in 1:nrow(formula_matrix2)) {
    print(f)
    fvector <- strsplit(unlist(formula_matrix2[f,1]), ",")[[1]]
    # create dag object syntax
    dag <- do.call(dagify, c(lapply(fvector, as.formula), additional_args))
    model <- full_matrix2$model[f]
    full_matrix_dags2[[model]] <- dag
    # extract adjustment sets
    mas <- adjustmentSets(dag, exposure = "Xtest", outcome = "Y", effect = "direct")
    mas <- ifelse(mas == "list()", "none", capture.output(mas))
    full_matrix2$mas[f] <- paste(unlist(mas), collapse = ",")
}
```

### Exploration Plotting

```{r explot}
dag_plots <- list()

for (l in 1:length(full_matrix_dags2)){
    p <- paste0("p",l)
    assign(p, 
           plot(full_matrix_dags2[[l]]))        
    text(0.5,-1, paste0("MAS =",full_matrix2$mas[l]))
    text(0,1, paste0("Model ",l))
    # this command stores the name of the plot in the list but not the plot
    dag_plots <- c(dag_plots, p)
}
```




## Specific modelling problem (delete later)

```{r testing}


test_matrix <- as.data.frame(t(matrix(c("Y ~ Xtest + X2 + X3, X3 ~ X2 + X1, Xtest ~ X3 + X1", 1))))

colnames(test_matrix) <- c("formula", "model")

for (f in 1:nrow(test_matrix)) {
    print(f)
    fvector <- strsplit(unlist(test_matrix[f,1]), ",")[[1]]
    # create dag object syntax
    dag <- do.call(dagify, c(lapply(fvector, as.formula), additional_args))
    # extract adjustment sets
    mas <- adjustmentSets(dag, exposure = "Xtest", outcome = "Y", effect = "direct")
    mas <- ifelse(mas == "list()", "none", capture.output(mas))
    test_matrix$mas[f] <- paste(unlist(mas), collapse = ",")
}

```



