---
title: "02 Working Paper"
author: "Nate Breznau & Hung H.V. Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

## R Markdown

```{r setup}

source("00_Run_Theory.R")

library("QCA")

```

## Generate Model Multiverse

### Input variables

Generate causal matrix

```{r models}
inputs <- list(nodes=c("a","b","c","d"), timing=c(0,-1,-2,-2),
              types=c("otc","test","ctr","ctr"))

causal_matrix <- build_causal_matrix(inputs)
```

### Derive formulas


#### Add models to causal matrix

```{r xmodels}

causal_matrix <- as.data.frame(causal_matrix)

causal_matrix$model <- as.numeric(causal_matrix$model)

m4 <- max(causal_matrix$model, na.rm = T)+1
# Model 4
causal_matrix[nrow(causal_matrix)+1,] <- c("Y", "Y", "~", m4, 0, "otc", 0, "otc")
causal_matrix[nrow(causal_matrix)+1,] <- c("X4", "X4", "~", m4, 0.5, "ctr", 0.5, "ctr")
causal_matrix[nrow(causal_matrix)+1,] <- c("Xtest", "Xtest", "~", m4, -1, "test", -1, "test")
causal_matrix[nrow(causal_matrix)+1,] <- c("X2", "X2", "~", m4, -2, "ctr", -2, "ctr")
causal_matrix[nrow(causal_matrix)+1,] <- c("X1", "X1", "~", m4, -2, "ctr", -2, "ctr")
causal_matrix[nrow(causal_matrix)+1,] <- c("Y", "X4", "~", m4, 0, "otc", 0.5, "ctr")
causal_matrix[nrow(causal_matrix)+1,] <- c("Xtest", "X4", "~", m4, -1, "test", 0.5, "ctr")
causal_matrix[nrow(causal_matrix)+1,] <- c("Xtest", "Y", "~", m4, -1, "test", 0, "otc")
causal_matrix[nrow(causal_matrix)+1,] <- c("X2", "Y", "~", m4, -2, "ctr", 0, "otc")
causal_matrix[nrow(causal_matrix)+1,] <- c("X2", "Xtest", "~", m4, -2, "ctr", -1, "test")
causal_matrix[nrow(causal_matrix)+1,] <- c("X1", "Xtest", "~", m4, -2, "ctr", -1, "test")
causal_matrix[nrow(causal_matrix)+1,] <- c("X1", "X2", "", m4, -2, "ctr", -2, "ctr")

causal_matrix$model <- as.numeric(causal_matrix$model)

m5 <- max(causal_matrix$model, na.rm = T)+1
# Model 6
causal_matrix[nrow(causal_matrix)+1,] <- c("Y", "Y", "~", m5, 0, "otc", 0, "otc")
causal_matrix[nrow(causal_matrix)+1,] <- c("X3", "X3", "~", m5, -1.5, "ctr", -1.5, "ctr")
causal_matrix[nrow(causal_matrix)+1,] <- c("Xtest", "Xtest", "~", m5, -1, "test", -1, "test")
causal_matrix[nrow(causal_matrix)+1,] <- c("X2", "X2", "~", m5, -2, "ctr", -2, "ctr")
causal_matrix[nrow(causal_matrix)+1,] <- c("X1", "X1", "~", m5, -2, "ctr", -2, "ctr")
causal_matrix[nrow(causal_matrix)+1,] <- c("X2", "X3", "~", m5, -2, "ctr", -1.5, "otc")
causal_matrix[nrow(causal_matrix)+1,] <- c("Xtest", "Y", "~", m5, -1, "test", 0, "otc")
causal_matrix[nrow(causal_matrix)+1,] <- c("X2", "Y", "~", m5, -2, "test", 0, "otc")
causal_matrix[nrow(causal_matrix)+1,] <- c("X2", "Xtest", "~", m5, -2, "ctr", -1, "test")
causal_matrix[nrow(causal_matrix)+1,] <- c("X1", "Xtest", "~", m5, -2, "ctr", -1, "test")
causal_matrix[nrow(causal_matrix)+1,] <- c("X3", "Xtest", "~", m5, -1.5, "test", 0, "ctr")

```

#### Generate formula matrix

```{r formula_matrix}

formula_matrix <- build_formula_matrix(causal_matrix)

```


### Derive plots

#### Update Model Numbers



#### Generate DAG formulas

```{r gen_dags}

dag_matrix <- build_dag_matrix(formula_matrix, add_nodes = 2)

```

#### Generate Plots

Only need to save models 1-6 for the paper

```{r gen_plots, warning = F}
dag_plots <- plot_dag_matrix(dag_matrix, save_plots = TRUE, choose_plots = c(2,6,8,18,33,34), title_labels = c(3,1,5,2,4,6))


```

#### DAG Plots

## Finding Multi MAS DAGs

Using our multiverse generator we search for DAGs that will not easily follow our classification rules

### Generate Models

#### Different Configurations

We will develop an easier method, but for now calculate different numbers of nodes in models with separate inputs and then combine them in a single causal matrix.

```{r searching_dags}
inputs1 <- list(nodes=c("a","b","c","d","e"),
                timing=c(0,-1,-2,-3,-3),
                types=c("otc","test","ctr","ctr", "ctr"))

inputs2 <- list(nodes=c("a","b","c","d"),
                timing=c(0,-1,-3,-3),
                types=c("otc","test","ctr","ctr"))

inputs3 <- list(nodes=c("a","b","c", "d"),
                timing=c(0,-1,-3, 1),
                types=c("otc","test","ctr", "ctr"))

causal_matrix1 <- build_causal_matrix(inputs1)
causal_matrix2 <- build_causal_matrix(inputs2)
causal_matrix3 <- build_causal_matrix(inputs3)

# by default inputs3 makes "d" = X2, change to X4
causal_matrix3 <- causal_matrix3 %>%
    mutate(from = gsub("X2", "X4", from),
           to = gsub("X2", "X4", to),
           direction = ifelse(is.na(direction),"~",direction))


lf1 <- length(causal_matrix1$model)
lf2 <- length(causal_matrix2$model)

causal_matrix2 <- causal_matrix2 %>%
    mutate(model = model + lf1)

causal_matrix3 <- causal_matrix3 %>%
    mutate(model = model + lf1 + lf2)
    

formula_matrix1 <- build_formula_matrix(causal_matrix1)
formula_matrix2 <- build_formula_matrix(causal_matrix2)
formula_matrix3 <- build_formula_matrix(causal_matrix3)

formula_matrix2 <- formula_matrix2 %>%
    mutate(model = model + lf1)

formula_matrix3 <- formula_matrix3 %>%
    mutate(model = model + lf1 + lf2)

formula_matrix_f <- formula_matrix1 %>%
    rbind(formula_matrix2) %>%
    rbind(formula_matrix3)

causal_matrix_f <- causal_matrix1 %>%
    rbind(causal_matrix2) %>%
    rbind(causal_matrix3)



```


### Set Logic Comparison

Starting with causal matrix
1. Identify each component (node / edge) and make categorical as yes/no 1/0
2. Select reference model 
3. Assign test_compatible and full_model_compatible



#### Identify components

```{r set_logic}
causal_matrix_transf <- causal_matrix_f %>%
    mutate(not_component = ifelse(direction == "", 1, 0),
           direction = ifelse(direction == "", "~", direction),
           component_link = ifelse(direction == "~~", "corr",
                            ifelse(direction == "~", "cause", NA)),
           component = paste(from, to, sep = "_"),
           component = ifelse(from == to, from, component),
           component_type = ifelse(from == to, "node", "edge"))
```


### Transform into sets

```{r set_build}

for (v in causal_matrix_transf$component) {
causal_matrix_transf[,paste0(v)] <- ifelse(causal_matrix_transf$component == v & causal_matrix_transf$not_component == 0, 1, 0)
}

causal_matrix_transf <- subset(causal_matrix_transf, select = -c(from, to, direction, timing_from, type_from, timing_to, type_to, component, not_component, component_link, component_type))

causal_matrix_transf <- causal_matrix_transf %>%
    group_by(model) %>%
    summarise_all(max, na.rm = T)


```

#### Add MAS

Use MAS as an example here, but in the future we will use test_compatible and test_model_compatible

```{r formula_mas}
mas <- subset(formula_matrix_f, select = c(model, mas))

causal_matrix_transf <- causal_matrix_transf %>%
    left_join(mas, by = "model")

# make MAS = "{X1}" the outcome, this fits with the X4 models
causal_matrix_transf <- causal_matrix_transf %>%
    mutate(outcome = ifelse(mas == "{ X1 }", 1, 0)) %>%
    select(-c(model,mas))
```


#### QCA

Use set logic to identify the MAS outcome { X2 }

Need to use a case study that has models with excluded nodes, and a confounder to see the full potential. 

The next step is to use test_compare and test_model_compare as the outcome

```{r qca}
test <- truthTable(as.data.frame(causal_matrix_transf), "outcome")

pof(setms = select(as.data.frame(causal_matrix_transf), -outcome), outcome = "outcome", data = as.data.frame(causal_matrix_transf))
```

