---
title: "02 Working Paper"
author: "Nate Breznau & Hung H.V. Nguyen"
date: "`r Sys.Date()`"
output: html_document
---

## R Markdown

```{r setup}

library('tictoc')
library('QCA')
library('theoRy') 

```

## Generate Model Multiverse

### Generate causal and formula matrix

Generating a 6 variable model multiverse takes about 3.5 hours. It will likely fail under 36 GB of RAM. Therefore, we save the results for the user and they can call them in the next code chunk.

```{r six_var_comp}
nodes <- c("a","b","c","d","e","f")
timing <- c(0,-1,-2,-3,-3, 1)
types <- c("otc","test","ctr","ctr", "ctr","ctr")
user_mods <- c("a ~ b + e; b ~ e + d", # Model 1
               "a ~ b + d; b ~ e + d", # Model 2
               "a ~ b + e + d; b ~ e + d", # Model 3
               "a ~ b + e; b ~ e + d; f ~ a + b", # Model 4
               "a ~ b + e; b ~ e; d ~ d", # Model 5
               "a ~ b + e; b ~ d + c + e; c ~ e" # Model 6
               )

tic()
ls_theory <- theoRy(nodes = nodes, timing = timing, types=types, include_subsets = TRUE, user_mods=user_mods)
toc()

# Demonstration: If we wanted to add models and/or model numbers after

ls_theory <- find_add_models(ls_theory = ls_theory, on_ls = TRUE, user_mods = c("a ~ b + e; b ~ e + d", # Model 1
               "a ~ b + d; b ~ e + d", # Model 2
               "a ~ b + e + d; b ~ e + d", # Model 3
               "a ~ b + e; b ~ e + d; f ~ a + b", # Model 4
               "a ~ b + e; b ~ e; d ~ d", # Model 5
               "a ~ b + e; b ~ d + c + e; c ~ e" # Model 6
               ), assert_mod_num = c(1,2,3,4,5,6))
```

### Load Matricies

```{r loadm}
saveRDS(ls_theory, here::here("Results", "ls_theory.RDS"))

ls_theory <- readRDS(here:::here("Results", "ls_theory.RDS"))
```

### Table 2

```{r t2}
Table_2 <- subset(ls_theory$causal_matrix, model %in% c(1,3,4,5), select = c(from, to, timing_from, timing_to, model))

write.csv(Table_2, here::here("Results", "Table_2.csv"), row.names = F)
```

## 1. A Potential Collider

### Compare Results

```{r cmp}
cmp_matrix <- add_compatible(ls_theory$formula_matrix, ref_mod = 1)

saveRDS(cmp_matrix, here::here("Results", "cmp_matrix.RDS"))

cmp_matrix <- readRDS(here::here("Results", "cmp_matrix.RDS"))

```

### Generate Plots. Table 1

Only need to save models 1-6 for the paper

```{r gen_plots, warning = F}
dag_plots <- plot_dag(ls_theory, choose_plots=c(1,2,3,4,5,6), save_path="Results/")

for(plot in dag_plots){
    print(plot)
}
```

## QCA

### Build Set Matrix

Here we focus on test_compatible as the outcome (= 1)

We use cmp_matrix to get this information

```{r build_set_matrix}
set_matrix <- build_set_matrix(ls_theory$causal_matrix, cmp_matrix = cmp_matrix, outcome_var = "test_compatible")

set_matrix_full_comp <-  build_set_matrix(ls_theory$causal_matrix, cmp_matrix = cmp_matrix, outcome_var= "full_model_compatible")

# think about whether we want model numbers in or not, just remove by hand here

# it makes no sense to have Xtest_Y because it is in every model

set_matrix_c <- dplyr::select(set_matrix, -c(model, Xtest_Y))
```




### Single Components

This can help us understand the role of single components. However, for most causal models a single component is rarely alone so important. For example a node often requires specific edges before it becomes important (e.g., becomes a confounder or collider)

```{r qca}
#set_truth <- truthTable(set_matrix_c, "outcome")

set_pof <- pof(setms = select(set_matrix, -c(model,outcome, Xtest_Y)), outcome = "outcome", data = set_matrix)

# try option relation = "sufficiency"

#inclN = inclusion = what percent of all models have outcome (Y = 1) and this component (X = 1)
#covN = necessity coverage = what percent of all models with this component (X = 1) have the outcome (Y = 1)
#RoN = relevence of necessity = is a necessary condition trivial

set_pof
```


### Multiple Component Analyses

#### Minimize

Looking only positive outcomes (outcome = 1/MAS = {X1}), this procedure tries to find minimal sets of variables that are present in all positive cases, regardless of if these sets are present in outcome = 0 cases.

```{r minimize}

min_results <- minimize(set_matrix_c, details = TRUE, outcome = "outcome")

```


#### Subsets

*inclN* = inclusion = what percent of all models have outcome (Y = 1) and this component (X = 1)
*covN* = necessity coverage = what percent of all models with this component (X = 1) have the outcome (Y = 1)
*RoN* = relevence of necessity = is a necessary condition trivial
```{r qca}
s_subset <- superSubset(set_matrix_c, outcome = "outcome")

s_subset
```
### Comparing the same # of variables

We have included all subsets. So these models have different variables. We might want to know about all causal models with a fixed set of variables. 

#### Identify vars

```{r var_find}
# identify models by their variables
subset_models <- as.data.frame(cmp_matrix) %>%
    mutate(X1 = ifelse(grepl("X1", unq_nodes), 1, 0),
           X2 = ifelse(grepl("X2", unq_nodes), 1, 0),
           X3 = ifelse(grepl("X3", unq_nodes), 1, 0),
           X4 = ifelse(grepl("X4", unq_nodes), 1, 0)) %>%
    subset(select = c(model, X1, X2, X3, X4))


```

#### Subset list

```{r subs}
models_X1X2X4 <- subset(subset_models, X1 == 1 & X2 == 1 & X3 == 0 & X4 == 1, select = model)$model

set_matrix_X1X2X4 <- subset(set_matrix, model %in% models_X1X2X4)
# flip to not_compatible

set_matrix_full_comp_X1X2X4 <- subset(set_matrix_full_comp, model %in% models_X1X2X4)

set_matrix_X1X2X4 <- set_matrix_X1X2X4 %>%
    mutate(outcome = (outcome-1)*-1)

set_matrix_full_comp_X1X2X4 <- set_matrix_full_comp_X1X2X4 %>%
    mutate(outcome = (outcome-1)*-1)

```

#### Subset X1X2X4 Test Compatible

So far this keeps M1 as the comparison model. Keep that in mind. 

We flipped the outcome because we want to know which are not test_compatible


*inclN* = inclusion = what percent of all models have outcome (Y = 1) and this component (X = 1)
*covN* = necessity coverage = what percent of all models with this component (X = 1) have the outcome (Y = 1)
*RoN* = relevence of necessity = is a necessary condition trivial

```{r subX1X2X4}
s_subset_X1X2X4 <- superSubset(dplyr::select(set_matrix_X1X2X4, -c(model, Xtest_Y)), outcome = "outcome")

s_subset_X1X2X4
```
Of all models with X1, X2 and X4, `r round(mean(set_matrix_X1X2X4$outcome)*100,1)`% are not test compatible with M1. If all models were equally plausible, a fair majority would not be test compatible with M1. 

### Subset X1X2X4 full model compatible

Again the inverse. So not full model compatible. 


```{r subfullx1x2x4}
s_subset_full_X1X2X4 <- superSubset(dplyr::select(set_matrix_full_comp_X1X2X4, -c(model, Xtest_Y)), outcome = "outcome")

s_subset_full_X1X2X4
```

None are compatible. Thats because our model does not have X4. To simulate this we need to compare a version of M1 that has a solo X4 (no paths to it) and compare that at full_model_compatible.

```{r next}

```

