---
title: "theoRy: Pipeline and Visualization Guide"
output:
  html_document:
    df_print: paged
vignette: >
  %\VignetteIndexEntry{theoRy: Pipeline and Visualization Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Setup

Install theoRy from GitHub and load the package. The visualization functions
require `ggplot2` and `ggdag`, which are already declared as imports.

```{r}
# install.packages("remotes")
# remotes::install_github("https://github.com/hungnguyen167/theoRy")
library(theoRy)
library(ggplot2)
```

Start the Python backend engine. This launches a FastAPI server that handles
registry building, state expansion, dyadic comparison, Delta-U computation,
simulation, and ghost cluster detection. Start it once in every R session that
uses the modern API and stop it when finished.

```{r}
start_theory_engine()
```

The backend also calls R through `rpy2`: Dagitty computes minimal adjustment
sets. The optional `causaleffect` dependency remains available for general-ID
callers, but it does not determine `identified_compatible`. After
upgrading theoRy, restart the engine because in-memory sessions from the
previous dyad schema are not compatible with the current API.

## Compatibility Contract

The modern API exposes exactly three pairwise metrics:

| Metric | Meaning |
|---|---|
| `similarity_rate` | Numeric component-level structural agreement. An edge applicable in only one model is one disagreement and one repair; an edge inapplicable in both is ignored. |
| `mas_compatible` | Both models have a robust shared minimal adjustment strategy for the same total-effect query. |
| `identified_compatible` | Both models satisfy the fixed-direct exposure-to-outcome complete-conditioning d-separation predicate and have exactly equal node-presence sets (all declared present nodes except exposure/outcome). The predicate removes only the mandatory direct edge and conditions on every other declared node, including mediators, confounders, and colliders; bidirected paths remain. Partial models require complete nonempty coverage and all valid descendants true; any false descendant is false and incomplete/empty coverage is unavailable. |

For MAS output, `[[]]` represents one valid empty adjustment set `{}`, whereas
`[]` means no valid adjustment set. Empty-set/empty-set can be compatible;
no-set/no-set is not. An unavailable profile becomes `NA` in R and is never
treated as a neutral score.

The causal metrics require both `exposure` and `outcome`, and causal queries
require a fixed causal `exposure -> outcome` registry edge applicable in every
queried model. Unknown edges are
evaluated through valid fully resolved descendants already in the multiverse,
and a positive necessary result requires complete descendant coverage.
Synthetic causal simulations can add support-only descendants to prove that
coverage, but those descendants are excluded from the analysis-model dyads,
rankings, aggregates, and clusters. Seeded runs require the supplied corpus to
be completion-closed.

## Quick Start: One-Call Workflow

`analyze_theory()` runs the entire five-step pipeline with a single call.
Provide variable names, timing values, and optional exposure/outcome and
prior model. Set `plot = TRUE` to generate standard visualizations.

```{r}
result <- analyze_theory(
  nodes = c("X1", "X2", "Y"),
  timing = c(1, 2, 3),
  exposure = "X1",
  outcome = "Y",
  prior_model = "M0001",
  mode = "exhaustive",
  plot = TRUE
)
```

The returned list contains every intermediate output:

```{r}
names(result)
cat(result$summary, sep = "\n")
```

Print the generated plots:

```{r}
for (p in result$plots) print(p)
```

When done, stop the engine:

```{r}
stop_theory_engine()
```

## Step-by-Step Pipeline

Each step can also be called individually for finer control.

### Step 1: Build the component registry

```{r}
registry <- build_component_registry(
  nodes = c("X1", "X2", "Y"),
  timing = c(1, 2, 3),
  exposure = "X1",
  outcome = "Y"
)
head(registry)
```

### Step 2: Expand model states

```{r}
# Exhaustive: enumerate all valid edge-status combinations
# Include all three statuses so both crux directions are resolution-closed.
states <- expand_model_states(registry, mode = "exhaustive",
  edge_statuses = c("causal", "unknown", "non-causal"))

# Or sampled: draw a random subset (kept separate because sampled states do
# not by themselves prove completion closure for causal metrics)
sampled_states <- expand_model_states(registry, mode = "sampled",
  n_models = 200, seed = 42)
```

### Step 3: Build the dyad matrix

```{r}
# Basic: structural metrics only
dyads <- build_dyad_matrix(registry, states, mode = "basic")

# Full is a computation/transport mode: include MAS and fixed-direct profiles
dyads <- build_dyad_matrix(registry, states, mode = "full",
  exposure = "X1", outcome = "Y")
head(dyads)
```

### Step 4: Compute Delta-U crux rankings

By default, Delta-U uses **marginal crux**: each applicable edge that is still
`unknown` in at least one model is ranked by evaluating both its causal and
non-causal resolutions against the unchanged baseline. The **global crux** mode
ranks each non-preset edge by forcing all of
that edge to causal and then to non-causal across every model where the edge is
applicable, including models where it is unknown or has the opposite resolved
status. Both directions are evaluated automatically.

```{r}
rankings <- compute_delta_u(
  dyads,
  top_k = 10,
  compatibility_metric = "identified_compatible"
)
head(rankings)

# Global edge ranking (both forcing directions are evaluated automatically)
global_crux <- compute_delta_u(
  dyads,
  crux_mode = "global",
  top_k = 10
)
global_crux
```

Both modes remap models to existing multiverse states and reuse their dyad
records; they never mutate claims or recompute causal profiles. The original
baseline model and dyad counts remain unchanged, with one explicit exception:
a hypothetical **directed causal** branch excludes timing-ineligible slots
whose source or target timing is missing, or whose source timing is greater
than or equal to target timing. Marginal mode applies this only among unknown
applicable instances; global mode applies it across all applicable models.
Non-causal and bidirected branches do not timing-prune, and no edge/component
is removed. For example, if a source is at time 2 and a flexible target may be
at time 1 or 3, only the time-1 causal slot is pruned. The result reports
per-direction retained `post_model_count_*` and `post_dyad_count_*` fields,
along with timing-pruned model IDs/counts and `insufficient_post_models_*`
flags; baseline `model_count` and `dyad_count` remain the original full
counts. Post compatibility is compared with that full baseline, and
improved/worsened counts cover retained aligned pairs. Marginal ranking still
requires a resolution-closed multiverse (e.g. exhaustive expansion) for every
retained remapping.

### Step 5: Detect ghost clusters

```{r}
ghost_result <- detect_ghost_clusters(
  dyads,
  prior_model = "M0001",
  score_field = "identified_compatible"
)
ghost_result$ghost_clusters
```

## Visualization Gallery

### Dyad similarity heatmap

```{r}
p <- plot_dyad_heatmap(dyads)
print(p)

# Causal compatibility heatmap
plot_dyad_heatmap(dyads, score_field = "mas_compatible")
```

### Lynchpin ranking bar chart

```{r}
plot_lynchpin_ranking(rankings)

# Show only top 5
plot_lynchpin_ranking(rankings, top_n = 5)
```

### UMAP cluster embedding

```{r}
plot_cluster_embedding(ghost_result)

# With cluster labels and prior model highlight
plot_cluster_embedding(ghost_result,
  show_labels = TRUE,
  highlight_prior = TRUE,
  prior_model_id = "M0001")
```

### Compatibility timeline (phase transition)

```{r}
lynchpin_sim <- run_simulation("lynchpin_of_certainty", n_models = 200)
plot_compatibility_timeline(lynchpin_sim)
```

### DAG models

```{r}
dag_plots <- plot_dag_models(registry, states,
  model_ids = c("M0001", "M0002"))
for (p in dag_plots) print(p)

# With minimal adjustment set annotation
dag_plots <- plot_dag_models(registry, states,
  model_ids = c("M0001"), show_mas = TRUE)
print(dag_plots[[1]])
```

## Simulation Scenarios

theoRy includes three proof-of-concept simulations. Crux of Certainty and Ghost
Discovery use the canonical `compatibility_metric` selector with one of
`"similarity_rate"`, `"mas_compatible"`, or `"identified_compatible"`.
Consensus Illusion requires one of the two causal metrics. The selected metric
drives the aggregate, shift, or clustering rather than only changing a label.
Use `plot_simulation()` to visualize the results.

### Consensus Illusion

High mean structural similarity masking a lower causal compatibility rate
across resolved and partial models.

```{r}
consensus <- run_simulation_consensus(
  n_models = 100,
  compatibility_metric = "mas_compatible"
)
plots <- plot_simulation(consensus)
for (p in plots) print(p)
```

Generated Consensus Illusion designs infer exposure `X1` and outcome `Y`. Their
canonical results report `mean_similarity_rate`, the selected
`compatibility_rate`, `consensus_illusion_gap`, resolved and partial model
counts, design, and diagnostics.

### Crux of Certainty

A single component resolution triggering a phase transition.

```{r}
lynchpin <- run_simulation(
  "crux_of_certainty",
  n_models = 200,
  compatibility_metric = "identified_compatible",
  exposure = "X1",
  outcome = "X2"
)
plots <- plot_simulation(lynchpin)
for (p in plots) print(p)
```

### Ghost Discovery

Discovering ghost clusters biased away from the user prior.

```{r}
ghost <- run_simulation(
  "ghost_discovery",
  n_models = 150,
  compatibility_metric = "similarity_rate"
)
plots <- plot_simulation(ghost)
for (p in plots) print(p)
```

Structural similarity is the default and recommended general signal for Ghost
Discovery. Boolean causal profiles are accepted only when all dyads are
available. The backend checks `metric_unique_values`, all-compatible and
all-incompatible flags, and profile variance; a no-variation matrix returns a
degenerate/no-cluster result rather than arbitrary UMAP output.

Simulation values are measured without clamping baseline or post-resolution
compatibility to acceptance bounds. Synthetic structural scenarios may retry a
seed when their calibrated threshold is missed. Symbolic simulations support
`similarity_rate` only because they do not materialize the fixed-direct
complete-conditioning profiles used by `identified_compatible`.

Simulations currently support directed components only, so
`include_bidirectional = TRUE` is rejected in both concrete and symbolic mode.
This restriction does not apply to general non-simulation APIs.

## Interpreting Results

- **Dyad heatmap**: Each cell shows the similarity (or compatibility) between
  two theories. Blue = high, red = low. Diagonal cells are self-pairs (NA).
  Clusters of blue cells indicate groups of similar theories.

- **Lynchpin ranking**: Horizontal bars show Delta-U scores. Taller blue bars
  indicate components that, if resolved as causal, would maximally increase
  global compatibility. Red bars indicate components best resolved as
  non-causal. Grey bars mark components whose resolution does not change
  global compatibility.

- **Cluster embedding**: UMAP projection of compatibility profiles. Red points
  are ghost clusters (internally consistent but prior-invisible). Blue points
  are mainstream (aligned with prior). Grey points are noise.

- **Compatibility timeline**: Shows how global compatibility changes as
  components are resolved step by step. A sharp jump indicates a phase
  transition.

- **DAG plots**: Each panel shows one theory's causal structure. Solid arrows
  are claimed causal relationships. Dashed gray arrows are unknown
  relationships. Non-causal edges are omitted.

```{r}
stop_theory_engine()
```
